---
title: "Cats of Monterey Bay, CA"
date: "`r format( Sys.Date(), '%B %d, %Y' )`"
output:
  html_document:
    theme: cosmo
---

```{r initializR, include = FALSE}

knitr::opts_chunk$set( "echo" = FALSE, 
                       "message" = FALSE, 
                       "warning" = FALSE )

```

```{r pkgR}

library( tidyverse )
library( lubridate )
library( leaflet )
library( sp )

```

```{r impoRt}

# Read in the raw data.
cats <- read.csv( file = "data/coastalCAcats.csv", 
                  header = TRUE, 
                  sep = "," )

# Consider only complete observations of feral (vs. wild) cats.
cats.complete <- cats[ complete.cases( cats ), ]
cats.complete <- cats.complete %>% 
  filter( type == "managed" ) %>% 
  dplyr::select( -type )


# Teach R to recognize dates.
cats.complete$date <- mdy( cats.complete$date )

# Derive a variable for Julian date.
cats.complete$Jday <- yday( x = cats.complete$date )

# Homogenize titer levels.
cats.complete$toxo.titer <- gsub( pattern = "<40", 
                                  replacement = "20", 
                                  x = cats.complete$toxo.titer )
cats.complete$toxo.titer <- gsub( pattern = "[^0-9]", 
                                  replacement = "", 
                                  x = cats.complete$toxo.titer )

# Correct likely errors in data entry. (Reviewed with E. VanWormer)
cats.complete$toxo.titer[ which( cats.complete$toxo.titer == "5180" ) ] <- "5120"
cats.complete$toxo.titer[ which( cats.complete$toxo.titer == "20240" ) ] <- "10240"

# Convert titers to numeric values.
cats.complete$toxo.titer <- as.numeric( cats.complete$toxo.titer )

# Correct infection-positive responses for different titer levels.
cats.complete$toxo40 <- ifelse( test = cats.complete$toxo.titer >= 40, 
                                yes = 1, 
                                no = 0 )
cats.complete$toxo160 <- ifelse( test = cats.complete$toxo.titer >= 160, 
                                 yes = 1, 
                                 no = 0 )
cats.complete$toxo320 <- ifelse( test = cats.complete$toxo.titer >= 320, 
                                 yes = 1, 
                                 no = 0 )

# Order age covariates.
cats.complete$age <- factor( x = cats.complete$age, 
                             levels = c( "< 6 months", "6-12 months", "> 12 months" ), 
                             ordered = TRUE )

cats.complete$age.2 <- factor( x = cats.complete$age.2, 
                               levels = c( "juvenile", "adult" ), 
                               ordered = TRUE )

# Data transformations
cats.complete$otoxo.titer <- as.ordered( cats.complete$toxo.titer )
cats.complete$fcolony <- as.factor( cats.complete$colony )

# Extract subset of data for which clinical infection (as measured by toxo titer) has been established.
cats.infect <- cats.complete[ which( as.logical( cats.complete$toxo160 ) ), ]

# Extract colony size data for summarization.
cats.colonies <- cats.complete %>% 
  dplyr::select( contains( "colony" ), -colony ) %>% 
  distinct()

```

###**C A T\ \ \ M A P !**

```{r mapR}

# Teach R to recognize longitude and latitude.
catsLocs <- cats %>% dplyr::select( c( longitude, latitude ) )
inds <- which( complete.cases( catsLocs ) )
catsLocs <- catsLocs[ inds, ]

# Re-create data as spatial.
catsSPDF <- SpatialPointsDataFrame( coords = catsLocs, 
                                    data = cats[ inds, ], 
                                    proj4string = CRS( "+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0" ) )


# Make a color palette for infection status representation.
pal <- colorFactor( c( "navy", "red" ), domain = c( 0, 1 ) )
aPal <- colorFactor( c( "yellow", "orange", "red" ), domain = c( "< 6 months", "6-12 months", "> 12 months" ) )

toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~pal( toxo160 ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

toxoMap

```

Each point is an individual cat, color-coded blue for *Toxoplasma*-negative cats (titer <1:160) and red for positive exposure status.

You can also do something with clusters, like this:

```{r clustR}

clusterCat <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~pal( toxo160 ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

clusterCat

```

Zooming and panning in this is pretty neat. I'd like to figure out a way to control color, etc. (for instance, coloring on a continuous gradient based on proportion infected, average titer, or some other metric), but I haven't really dug into that yet.

-----

###**C O L O N Y\ \ \ M A P !**

```{r colonyMapR}

colonies <- cats %>%
  group_by( colony ) %>%
  summarize( size = n(),
             toxo = sum( toxo160 ) * 100 / n(),
             latitude = mean( latitude ),
             longitude = mean( longitude ) )
colonies <- colonies[ complete.cases( colonies ), ]

coloniesLocs <- colonies %>% dplyr::select( c( longitude, latitude ) )

coloniesSPDF <- SpatialPointsDataFrame( coords = coloniesLocs,
                                        data = colonies,
                                        proj4string = CRS( "+init=epsg:4326 +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0" ) )

qpal <- colorNumeric( palette = "YlOrRd", domain = coloniesSPDF$toxo )

colonyMap <- leaflet( data = coloniesSPDF ) %>%
  addTiles() %>%
  addCircleMarkers( lng = ~longitude,
                    lat = ~latitude,
                    radius = ~size,
                    color = ~qpal( toxo ),
                    stroke = FALSE,
                    label = ~paste( colony ), 
                    labelOptions = labelOptions( noHide = TRUE, 
                                                 textOnly = TRUE ), 
                    fillOpacity = 1 )

colonyMap

```

Here's another map looking at colonies only. The size of the marker corresponds to the number of cats tested from the colony, and the color on the yellow--orange--red palette indicates the proportion with a history of exposure to *Toxoplasma*.

-----

###**E X P L O R E !**

**n** = `r nrow(cats)` cats.

####***Toxoplasma***

```{r toxo160}

toxoMap

```

Let's first revisit cases of toxoplasmosis. Red points are *Toxoplasma*-positive cats (titer ≥1:160; **n** = `r sum(cats$toxo160,na.rm=TRUE)`).

####**FIV**

```{r fiv}

fivMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~pal( fiv ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fivMap

```

Okay. Now here's a look at FIV exposure (**n** = `r sum(cats$fiv,na.rm=TRUE)`).

####**FeLV**

```{r felv}

felvMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~pal( felv ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

felvMap

```

And FeLV infection (**n** = `r sum(cats$felv,na.rm=TRUE)`).

Now on to coinfections.

####**FIV--FeLV**

```{r fivfelv}

# Make a color palette for infection status representation.
palCo <- colorFactor( c( "navy", "navy", "navy", "red" ), domain = c( "0.0", "0.1", "1.0",  "1.1" ) )

fiv.felvMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo( interaction( fiv, felv ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fiv.felvMap

```

FIV--FeLV coexposure/infection (**n** = `r sum(cats$fiv==1&cats$felv==1,na.rm=TRUE)`).

####**FIV--_Toxoplasma_**

```{r fivtoxo}

fiv.toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo( interaction( fiv, toxo160 ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fiv.toxoMap

```

FIV--*Toxoplasma* coexposure (**n** = `r sum(cats$fiv==1&cats$toxo160==1,na.rm=TRUE)`).

####**FeLV--_Toxoplasma_**

```{r felvtoxo}

felv.toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo( interaction( felv, toxo160 ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

felv.toxoMap

```

FIV--*Toxoplasma* coexposure (**n** = `r sum(cats$felv==1&cats$toxo160==1,na.rm=TRUE)`).

####**FIV--FeLV--_Toxoplasma_**

```{r fivfelvtoxo}

# Make a color palette for infection status representation.
palCo2 <- colorFactor( c( "navy", "navy", "navy", "navy", "navy", "navy", "navy", "red" ), domain = c( "0.0.0", "0.0.1", "0.1.0", "0.1.1", "1.0.0", "1.0.1", "1.1.0", "1.1.1"  ) )


fiv.felv.toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo2( interaction( fiv, felv, toxo160 ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fiv.felv.toxoMap

```

FIV--*Toxoplasma* coexposure (**n** = `r sum(cats$fiv==1&cats$felv==1&cats$toxo160==1,na.rm=TRUE)`).

Those maps are pretty crowded by negative cases. Let's eliminate those...

####***Toxoplasma***

```{r toxo160}

toxoMap %>% addCircleMarkers( color = "red" )

```

Let's first revisit cases of toxoplasmosis. Red points are *Toxoplasma*-positive cats (titer ≥1:160; **n** = `r sum(cats$toxo160,na.rm=TRUE)`).

####**FIV**

```{r fiv}

fivMap <- leaflet( data = filter( catsSPDF, fiv == 1 ) ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~pal( fiv ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fivMap

```

Okay. Now here's a look at FIV exposure (**n** = `r sum(cats$fiv,na.rm=TRUE)`).

####**FeLV**

```{r felv}

felvMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~pal( felv ), 
                    stroke = FALSE, 
                    fillOpacity = 0.7 )

felvMap

```

And FeLV infection (**n** = `r sum(cats$felv,na.rm=TRUE)`).

Now on to coinfections.

####**FIV--FeLV**

```{r fivfelv}

# Make a color palette for infection status representation.
palCo <- colorFactor( c( "transparent", "transparent", "transparent", "red" ), domain = c( "0.0", "0.1", "1.0",  "1.1" ) )

fiv.felvMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo( interaction( fiv, felv ) ), 
                    stroke = FALSE, 
                    fillOpacity = 1 )

fiv.felvMap

```

FIV--FeLV coexposure/infection (**n** = `r sum(cats$fiv==1&cats$felv==1,na.rm=TRUE)`).

####**FIV--_Toxoplasma_**

```{r fivtoxo}

fiv.toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo( interaction( fiv, toxo160 ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fiv.toxoMap

```

FIV--*Toxoplasma* coexposure (**n** = `r sum(cats$fiv==1&cats$toxo160==1,na.rm=TRUE)`).

####**FeLV--_Toxoplasma_**

```{r felvtoxo}

felv.toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo( interaction( felv, toxo160 ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

felv.toxoMap

```

FIV--*Toxoplasma* coexposure (**n** = `r sum(cats$felv==1&cats$toxo160==1,na.rm=TRUE)`).

####**FIV--FeLV--_Toxoplasma_**

```{r fivfelvtoxo}

# Make a color palette for infection status representation.
palCo2 <- colorFactor( c( "navy", "navy", "navy", "navy", "navy", "navy", "navy", "red" ), domain = c( "0.0.0", "0.0.1", "0.1.0", "0.1.1", "1.0.0", "1.0.1", "1.1.0", "1.1.1"  ) )


fiv.felv.toxoMap <- leaflet( data = catsSPDF ) %>% 
  addTiles() %>% 
  addCircleMarkers( lng = ~longitude, 
                    lat = ~latitude, 
                    # clusterOptions = markerClusterOptions(), 
                    radius = 7, 
                    color = ~palCo2( interaction( fiv, felv, toxo160 ) ), 
                    stroke = FALSE, 
                    fillOpacity = 0.3 )

fiv.felv.toxoMap

```

FIV--*Toxoplasma* coexposure (**n** = `r sum(cats$fiv==1&cats$felv==1&cats$toxo160==1,na.rm=TRUE)`).